Synthetic Observations of Extremely High Velocity (EHV) SiO component in young molecular outflows

### 2013-11-26 PAPER/Paola_1/ms_ver2.tex
Added appendix to the paper to discuss the shock velocities

>>>>

\appendix
\section{Working surfaces and shocks}
In this section we describe the mathematical derivation of estimating
the speed of the working surface of the internal knot and shock speeds
associated with it. We will follow the derivation as given in
\citep{Raga:1990p16416,Raga:1992p16392}. Fig.~\ref{fig:interknot} shows a cartoon
representation of a single internal knot formed in the jet. 
The density and velocity upstream of the internal shock are
$\rho_{1}$ and $u_{1}$, while, the downstream values are $\rho_{2}$
and $u_{2}$ respectively. Note the velocities u$_1$ and u$_2$ are measured
in a stationary frame of reference of the star. The velocity of the working surface is
denoted by $v$, can be derived from the ram pressure balance between the upstream and
downstream flow,
\begin{equation}
\rho_1 (u_1 - v)^{2} = \rho_2 (v - u_2)^{2}.
\label{eq:ramPbal}
\end{equation}
This gives us the expression for $v$ as, 
\begin{equation}
v = \frac{u_1 + \zeta u_2}{1 + \zeta},
\label{eq:velws}
\end{equation} 
where, $$\zeta = \sqrt{\frac{\rho_{2}}{\rho_1}}$$.

\begin{figure}
\includegraphics[width=1\columnwidth]{\figpath/Internalknot.pdf}
\caption{Schematic diagram showing the structure of the internal
  working surface. The faster upstream material with velocity u$_1$
  catches the slower downstream material with velocity u$_2$ forming a
two shock structure (u$_2 >$ v $>$ u$_1$). The in-between gas is squeezed out sideways
forming the cocoon.}
\label{fig:interknot}
\end{figure}

The parameter $\zeta$ is obtained in principle from the two
dimensional snapshot of the jet density. In case of bow shock, the
downstream medium is the stationary ambient medium (u$_2$ = 0), thus,
eq.~\ref{eq:velws} will reduce to eq~\ref{eq:shockjetvel}.
However for the internal knots, the variation of density and
velocity for the reference run is shown in fig.~\ref{fig:1dvelrho}.
As can be clearly seen, the value of $\zeta \approx$ 1 for most of the
internal knots seen in our simulations. Thus, the velocity of the
working surface simplifies to, 
\begin{equation}
v = \frac{u_1 + u_2}{2}.
\label{eq:velwssimple}
\end{equation}

Additionally, the shock velocity (i.e, the velocity of the pre-shock
flow with respect to the velocity of the working surface) of the
upstream and downstream facing shocks have the same velocity given by \cite{Raga:1990p16416},
\begin{equation}
w = \frac{u_2 - u_1}{2} = 0.5\Delta u.
\label{eq:shkvelws}
\end{equation}

\begin{figure}
\includegraphics[width=1\columnwidth]{\figpath/jet_rhovel.pdf}
\caption{Variation of jet velocity (blue) and density (red) as a
  function of height close to the axis of the flow. The inset on the
  bottom left of the figure shows a zoomed variation of velocity and
  clearly marks the peak to peak difference between the upstream and
  downstream components.}
\label{fig:1dvelrho}
\end{figure}

The difference in the pre- and post-discontinuity velocities, $\Delta
u$ is marked clearly in the fig.~\ref{fig:1dvelrho}. Typically, the
difference in velocity values for our reference runs lie around 15-40 km s$^{-1}$. Thus, the
shock velocity $w$ lies around 10-20 km s$^{-1}$. Similar values are
also obtained for jets like HH 34 by \cite{Raga:1992p16392}.

The variation of SiO abundance used for the present model is rather
obtained from empirical values. In case of L1448, \cite{MartinPintado:1992p14309}
have shown that for extreme velocities (obtained from shift in the
spectra) of -40 to -50 km\,s$^{-1}$, the SiO abundance could be as
large as $6\times10^{-7}$, while, it slightly reduces to
$2\times10^{-7}$ for the velocity range of -60 to -50
km\,s$^{-1}$. For the same source, typical {\em{shock}} (or working
surface) velocities of the {\em{molecular bullets}} are estimated to
lie around 80-175 km\,s$^{-1}$ \cite{Dutrey:1997p11185}.
\\
{\color{red}
WE HAVE TWO OPTIONS:\\ 
1. Since the there is a empirical evidence with bulk gas motion. We
can use just the jet velocity obtained from simulations.\\
OR\\
2. We can try to connect the working surface velocities from
\cite{Dutrey:1997p11185} to abundance obtained by
\cite{MartinPintado:1992p14309} and there by use the velocity of the
working surface as an independent variable on the X axis of the
abundance plot.
}
